Intended Table of Contents Thesis

### 2013-11-18 Chapters/Chapter4/data_gathering.tex
Add caption titles to tables
Restructure Chapter 5 graphs of data sets

% !TEX root = ../../main.tex
\section{Data Gathering}\label{sec:method_data_gathering}
In this section we briefly discuss the different data gathering methods used for the change detection algorithms and experiments.
Section~\ref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
In Section~\ref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
Both sections refer to Chapters~\ref{Chapter5} and~\ref{Chapter6} for more details, respectively.

% ------
\subsection{Artificial data}\label{subsec:data_gathering_artificial}
In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the researches on which our method is based.
These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
Both construct a collection of one-dimensional time series data according to a second order \gls{ar} model:
\begin{equation}
  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t.
\end{equation}
In the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
Using this data set an objective quality measure over the change detection methods can be obtained and compared.
All the used data sets are listed and analyzed in Chapter~\ref{Chapter5}.

% ------
\subsection{Real-world data}\label{subsec:data_gathering_real_world}
In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
For our setup we record the activities of humans performed both in- and outdoor in an uncontrolled environment.
Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
Our method uses the signals from the accelerometer, gravity, and rotation sensors.
These time series data are used to detect change points.
A video recording from the performed activity is used to annotate the time series with real change points.
The discovered change point are compared with these annotated change points to give a subjective quality measure.
In Chapter~\ref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

@@

\begin{center}\begin{table}
  \begin{tabulary}{\textwidth}{|l|L|c|c|}
    \hline
    Metric & Description & Units of measure & Typical range \\
    \hline \hline
    Accelerometer & Acceleration force along each axis. & $m/s^2$ & $-20$ -- $20$ \\
    \hline
    Gravity & Force of gravity along each axi.s & $m/s^2$ & $-10$ -- $10$\\
    \hline
    Gyroscope & & $rad/s$ & $-15$ -- $15$\\
    \hline
    Light & Light sensitive sensor at the front of the phone. & & $0$ -- $10000$ \\
    \hline
    Linear acceleration & & $m/s^2$ & $-20$ -- $20$ \\
    \hline
    Magnetic field & Geomagnetic field strength along each axis. & $\mu T$ & $-60$ -- $60$ \\
    \hline
    Orientation & & & $-100$ -- $360$ \\
    \hline
    Rotation & & & $-1$ -- $1$\\
    \hline
  \end{tabulary}
  \caption{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
  \label{tab:recorded_metrics}
\end{table}\end{center}

>>>>

% !TEX root = ../../main.tex
\section{Data Gathering}\label{sec:method_data_gathering}
In this section we briefly discuss the different data gathering methods used for the change detection algorithms and experiments.
\Cref{subsec:data_gathering_artificial} reviews the artificial data sets we will use.
In \Cref{subsec:data_gathering_real_world} an overview of the real-world data sets used is provided.
Both sections refer to Chapters~\ref{Chapter5} and~\ref{Chapter6} for more details, respectively.

% ------
\subsection{Artificial data}\label{subsec:data_gathering_artificial}
In order to provide an objective comparison to other methods, we will use artificial data sets which are also used in the researches on which our method is based.
These are the data sets used by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change}.
Both construct a collection of one-dimensional time series data according to a second order \gls{ar} model:
\begin{equation}
  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t.
\end{equation}
In the different data series the mean and variance of the Gaussian random variable $\epsilon_t$ differs and changes at pre-determined change points.
Using this data set an objective quality measure over the change detection methods can be obtained and compared.
All the used data sets are listed and analyzed in \Cref{Chapter5}.

% ------
\subsection{Real-world data}\label{subsec:data_gathering_real_world}
In the second type of data sets we apply our method of change detection and temporal segmentation to real-world data sets.
For our setup we record the activities of humans performed both in- and outdoor in an uncontrolled environment.
Activities performed include sitting, standing, walking, running in a straight and curved line, and walking up- and downstairs.
Our method uses the signals from the accelerometer, gravity, and rotation sensors.
These time series data are used to detect change points.
A video recording from the performed activity is used to annotate the time series with real change points.
The discovered change point are compared with these annotated change points to give a subjective quality measure.
In \Cref{Chapter6} we give a detailed analysis of the performed activities and the recorded data sets.

@@

\begin{center}\begin{table}
  \begin{tabulary}{\textwidth}{|l|L|c|c|}
    \hline
    Metric & Description & Units of measure & Typical range \\
    \hline \hline
    Accelerometer & Acceleration force along each axis. & $m/s^2$ & $-20$ -- $20$ \\
    \hline
    Gravity & Force of gravity along each axi.s & $m/s^2$ & $-10$ -- $10$\\
    \hline
    Gyroscope & & $rad/s$ & $-15$ -- $15$\\
    \hline
    Light & Light sensitive sensor at the front of the phone. & & $0$ -- $10000$ \\
    \hline
    Linear acceleration & & $m/s^2$ & $-20$ -- $20$ \\
    \hline
    Magnetic field & Geomagnetic field strength along each axis. & $\mu T$ & $-60$ -- $60$ \\
    \hline
    Orientation & & & $-100$ -- $360$ \\
    \hline
    Rotation & & & $-1$ -- $1$\\
    \hline
  \end{tabulary}
  \caption[Measured metrics]{Measured metrics. The set of axis is always the triple (x, y, z) direction.}
  \label{tab:recorded_metrics}
\end{table}\end{center}

### 2013-11-18 Chapters/Chapter4/model_construction.tex
Restructure Chapter 5 graphs of data sets

The size of the initial window of data objects has a lower bound determined by the hyperparameter $C$ (Equation~\ref{eq:svdd_objective}).
Because of the equality constraint $\sum_{i=1}^n a_i x_i = 1$ and the box constraint $0 \le x_i \le C$, the number of objects in the working set must be at least $\ceil{\frac{1}{C}}$.
Thus the algorithm is initialized by selection the first $\ceil{\frac{1}{C}}$ objects for the working set.
In every loop of the algorithm the number of elements added are at least equal to the number of elements removed.
The optimality of the algorithm is proved by analysis of the \gls{kkt} conditions \cite{tax2003online}.

From experiments it shows that the (online) \gls{isvdd} method results in less false alarms than the static \gls{svdd}.
An explanation for this is that \gls{isvdd} follows the changing data distribution, \ie small changes over time.
Such changes are a drift in mean or increase in frequency.
The \gls{isvdd} algorithm continuously re-models the \gls{svm} representation.
This re-modeling of the \gls{svm} representation is beneficial for our purpose, since we are only interested in sudden changes between activities.
The following section will discuss the features extracted from the constructed \gls{oc-svm} model at every loop-step of the algorithm, before Section~\ref{sec:method_change_detection} describes how these features are used to detect change.

>>>>

The size of the initial window of data objects has a lower bound determined by the hyperparameter $C$ (\Cref{eq:svdd_objective}).
Because of the equality constraint $\sum_{i=1}^n a_i x_i = 1$ and the box constraint $0 \le x_i \le C$, the number of objects in the working set must be at least $\ceil{\frac{1}{C}}$.
Thus the algorithm is initialized by selection the first $\ceil{\frac{1}{C}}$ objects for the working set.
In every loop of the algorithm the number of elements added are at least equal to the number of elements removed.
The optimality of the algorithm is proved by analysis of the \gls{kkt} conditions \cite{tax2003online}.

From experiments it shows that the (online) \gls{isvdd} method results in less false alarms than the static \gls{svdd}.
An explanation for this is that \gls{isvdd} follows the changing data distribution, \ie small changes over time.
Such changes are a drift in mean or increase in frequency.
The \gls{isvdd} algorithm continuously re-models the \gls{svm} representation.
This re-modeling of the \gls{svm} representation is beneficial for our purpose, since we are only interested in sudden changes between activities.
The following section will discuss the features extracted from the constructed \gls{oc-svm} model at every loop-step of the algorithm, before \Cref{sec:method_change_detection} describes how these features are used to detect change.

### 2013-11-18 Chapters/Chapter4/model_features.tex
Restructure Chapter 5 graphs of data sets

The \gls{isvdd} algorithm creates a spherical \gls{oc-svm} representation of the working set at every step of the algorithm.
This model is obtained by the minimization of Equation~\ref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
We will use the the radius $R$ of the hypersphere as an indication of change.

In \cite{tax2002uniform} Tax and Duin provide an analysis of the error of the \gls{svdd} algorithm.
This error is based on
\begin{inparaenum}[\itshape 1\upshape)]
\item the fraction $f_{T-}$ of target objects that is rejected, and
\item the fraction $f_{O+}$ of outliers that is accepted.
\end{inparaenum}
Since in \gls{occ} situation typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the outliers are uniformly distributed around the target set.
To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O+}$, we should minimize the volume of the target data description (\ie the boundary of \gls{svdd}).
This is because the fraction of accepted outliers $f_{O+}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, \ie the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
This optimization will result in the modification of the radius $R$ of Equation~\ref{eq:svdd_objective} and affects the Lagrangian inequality~(\ref{eq:svdd_inequality}).

Since in our method the parameters $C$ and $\sigma$ are kept constant, and thereby also the fraction of target objects being rejected, the only free parameter is the radius $R$.
During the \gls{isvdd} algorithm the volume of the hypersphere is being minimized.
This means that only the radius of the sphere will change during the algorithm.
We will interpret changes in the radius $R$ with the same continuity assumptions as mentioned in Section~\ref{subsec:kernels}.
This means that when two objects (in this case radius sizes) are close to each other, their original objects are also close.
In other words, when the underlying distribution of time series is changed, and thus the characteristics of the data objects changes, the corresponding radius $R$ will also change.
Since in that case the data objects in the working set become more heterogeneous, the radius of the hypersphere will increase.
When, instead, the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.

>>>>

The \gls{isvdd} algorithm creates a spherical \gls{oc-svm} representation of the working set at every step of the algorithm.
This model is obtained by the minimization of \Cref{eq:svdd_objective}, which incorporates the radius $R$ of the sphere and the distances $\vectorsym{\xi}$ from the outliers to the boundary.
We will use the the radius $R$ of the hypersphere as an indication of change.

In \cite{tax2002uniform} Tax and Duin provide an analysis of the error of the \gls{svdd} algorithm.
This error is based on
\begin{inparaenum}[\itshape 1\upshape)]
\item the fraction $f_{T-}$ of target objects that is rejected, and
\item the fraction $f_{O+}$ of outliers that is accepted.
\end{inparaenum}
Since in \gls{occ} situation typically there are (almost) no examples of outlier objects, Tax and Duin construct a method to generate outliers based on the assumption that the outliers are uniformly distributed around the target set.
To minimize the error, calculated by the fractions $f_{T-}$ and $f_{O+}$, we should minimize the volume of the target data description (\ie the boundary of \gls{svdd}).
This is because the fraction of accepted outliers $f_{O+}$ is an estimate of the volume of the target data description, with respect to the volume of the outlier distribution.
Tax and Duin provide a method to optimize the parameters of the \gls{svdd} method, \ie the trade-off parameter $C$ and the \gls{rbf} kernel width $\sigma$.
This optimization will result in the modification of the radius $R$ of \Cref{eq:svdd_objective} and affects the Lagrangian inequality~(\ref{eq:svdd_inequality}).

Since in our method the parameters $C$ and $\sigma$ are kept constant, and thereby also the fraction of target objects being rejected, the only free parameter is the radius $R$.
During the \gls{isvdd} algorithm the volume of the hypersphere is being minimized.
This means that only the radius of the sphere will change during the algorithm.
We will interpret changes in the radius $R$ with the same continuity assumptions as mentioned in \Cref{subsec:kernels}.
This means that when two objects (in this case radius sizes) are close to each other, their original objects are also close.
In other words, when the underlying distribution of time series is changed, and thus the characteristics of the data objects changes, the corresponding radius $R$ will also change.
Since in that case the data objects in the working set become more heterogeneous, the radius of the hypersphere will increase.
When, instead, the data objects change from a heterogeneous set to a more homogeneous, we expect the radius to decrease in value, since the data objects are closer to each other in the feature space.

### 2013-11-18 Chapters/Chapter4/overview.tex
Add notes from gerard
Add caption titles to tables
Restructure Chapter 5 graphs of data sets

\begin{figure}
  \centering
    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup.pdf}
  \caption[Method setup]{Schematic representation of the change detection method.}
  \label{fig:method_overview}
\end{figure}

This section gives a description of the method used for the experiments and change detection mechanism.
First described is the method to process the gathered sensor data.
A schematic overview is given in figure~\ref{fig:method_overview} and shows the steps of the method.
A more detailed explanation of the ``Update model'' step follows.
This section is finalized with the experiments setup, annotation of data streams and quality measure.

\subsection{Change detection method}
As graphically represented in figure~\ref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.

>>>>

\begin{figure}
  \centering
    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/method_setup.pdf}
  \caption[Method setup]{Schematic representation of the change detection method. \TODO{Gerard noemt dit misschien ook goed voor bij de inleiding, als motivatie voor onderzoek. Overnemen?}}
  \label{fig:method_overview}
\end{figure}

This section gives a description of the method used for the experiments and change detection mechanism.
First described is the method to process the gathered sensor data.
A schematic overview is given in \Cref{fig:method_overview} and shows the steps of the method.
A more detailed explanation of the ``Update model'' step follows.
This section is finalized with the experiments setup, annotation of data streams and quality measure.

\subsection{Change detection method}
As graphically represented in \Cref{fig:method_overview}, the change detection method starts by processing the data from sensor, such as the accelerometer, magnetic orientation and rotation measures \footnote{Something about the origin of the streams; all from the same sensor or different sensors?}.

### 2013-11-18 Chapters/Chapter5/artificial_data_generation.tex
Restructure Chapter 5 graphs of data sets

% !TEX root = ../../main.tex
\section{Artificial Data}\label{sec:artificial_data}
\TODO{Use data from \cite{camci2010change,takeuchi2006unifying}}
In this section we present three data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci.eps}
  \caption[Jumping mean camci]{Jumping mean camci}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_thresholds.eps}
  \caption[Jumping mean camci thresholds]{Jumping mean camci, thresholds}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_mean_camci_ratios_10.eps}
  \caption[Jumping mean camci ratios]{Jumping mean camci ratios, last 10}
  \label{fig:jumping_mean_ratios}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance.eps}
  \caption[Jumping variance]{Jumping variance}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_thresholds.eps}
  \caption[Jumping variance thresholds]{Jumping variance, thresholds}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/jumping_variance_ratios_10.eps}
  \caption[Jumping variance ratios]{Jumping variance ratios, last 10}
\end{figure}

>>>>

% !TEX root = ../../main.tex
\section{Artifical data generation}\label{sec:artificial_data_generation}
In this section we present four data sets as used by \cite{camci2010change,takeuchi2006unifying} to provide for a objetive performance comparison.
All the data sets are modeled by a second-order \gls{ar}-model as in \Cref{eq:artificial_data_sets_model}, where $\epsilon_t$ is a Gaussian distribution modeling the noise with mean $0$ and variance $\sigma^2 = 1$, $a_1 = 0.6$ and $a_2 = -0.5$.
\begin{equation}\label{eq:artificial_data_sets_model}
  x_t = a_1 x_{t-1} + a_2 x_{t-2} + \epsilon_t,
\end{equation}
The length of $x_t$ is $10000$ and change points are generated at each $y \times 10000^\text{th}$ data point, with $y = (1, 2, 3, \dots 9)$.

Using this general \gls{ar}-model, we create four different data sets.
Each data set is characterized by a change in mean, variance, or both of the Gaussian term $\epsilon_t$:
\begin{enumerate}
  \item \textbf{Fixed increasing mean:} at each change point $y$ the mean is increased with $\Delta(y) = 5$.
  This set origins as the first data set of Camci~\cite{camci2010change}.
  See \Cref{fig:camci_fixed_increasing_mean,fig:camci_fixed_increasing_mean_thresholds}.

  \item \textbf{Reduced increasing mean:} at each change point $y$ the mean is increased with $\Delta(y) = 10 - y$.
  This set origins as the first data set of Takeuchi \etal \cite{takeuchi2006unifying}.
  See \Cref{fig:takeuchi_reduced_increasing_mean,fig:takeuchi_reduced_increasing_mean_thresholds}.

  \item \textbf{Reduced increasing mean, increasing variance:} at each change point $y$ the mean is increased with $\Delta(y) = 10 - y$ and the standard deviation is of $\epsilon_t$ is $0.1 / (0.01 + (10000 - x)/1000)$.
  This set origins as the second set of Camci~\cite{camci2010change}.
  See \Cref{fig:camci_reduced_increasing_mean_increasing_variance,fig:camci_reduced_increasing_mean_increasing_variance_thresholds}

  \item \textbf{Alternating variance:} at each change point $y$ the variance $\sigma^2$ is set to $9.0$ when $y$ is odd, and to $1.0$ otherwise.
  This set origins as the thirds sets from Camci~\cite{camci2010change} and Takeuchi \etal \cite{takeuchi2006unifying}.
  See \Cref{fig:camci_takeuchi_alternating_variance,fig:camci_takeuchi_alternating_variance_thresholds}.
\end{enumerate}

% =====

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean.eps}
  \caption[Fixed increasing mean]{Fixed increasing mean}
  \label{fig:camci_fixed_increasing_mean}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/camci_fixed_increasing_mean_thresholds.eps}
  \caption[Fixed increasing mean, thresholds]{Fixed increasing mean, thresholds}
  \label{fig:camci_fixed_increasing_mean_thresholds}
\end{figure}

% =====

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean.eps}
  \caption[Reduced increasing mean]{Reduced increasing mean}
  \label{fig:takeuchi_reduced_increasing_mean}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/takeuchi_reduced_increasing_mean_thresholds.eps}
  \caption[Reduced increasing mean, thresholds]{Reduced increasing mean, thresholds}
  \label{fig:takeuchi_reduced_increasing_mean_thresholds}
\end{figure}

% =====

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance.eps}
  \caption[Reduced increasing mean, increasing variance]{Reduced increasing mean, increasing variance.}
  \label{fig:camci_reduced_increasing_mean_increasing_variance}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/camci_reduced_increasing_mean_increasing_variance_thresholds.eps}
  \caption[Reduced increasing mean, increasing variance, thresholds]{Reduced increasing mean, increasing variance, thresholds (log scale)}
  \label{fig:camci_reduced_increasing_mean_increasing_variance_thresholds}
\end{figure}

% =====

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance.eps}
  \caption[Alternating variance]{Alternating variance}
  \label{fig:camci_takeuchi_alternating_variance}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1\textwidth]{./Figures/notes/camci_takeuchi_alternating_variance_thresholds.eps}
  \caption[Alternating variance, thresholds]{Alternating variance, thresholds}
  \label{fig:camci_takeuchi_alternating_variance_thresholds}
\end{figure}

### 2013-11-18 intended_contents.tex
Restructure Chapter 5 graphs of data sets

\section{Introduction}
\label{sec:Introduction}
This section introduces the subject, outlines the global goal of the thesis and explains the problem.
The main research-questions are introduced.
In the subsections the main components (global, not as an implementation) are introduced.
Outline of the thesis is presented.
Chapter~\ref{sec:Concepts} introduces the big parts of the puzzle.
Chapter~\ref{sec:Setup} gives details about the algorithms used.
Chapter~\ref{sec:Implementation} will give the implementation details.
Chapter~\ref{sec:Experiments} will show how data is gathered and shows the data.
Chapter~\ref{sec:Discussion} will discuss and relate the results
Chapter~\ref{sec:Conclusion} will conclude and give final remarks.

@@

\section{Concepts}
\label{sec:Concepts}
This chapter will introduce the abstract concepts which will be used.
There will be some formulas, but minimal implementations.
The focus is on the meaning of the terms and what they do and how they do it.
Real implementations will be discussed in chapter~\ref{sec:Setup}.

>>>>

\section{Introduction}
\label{sec:Introduction}
This section introduces the subject, outlines the global goal of the thesis and explains the problem.
The main research-questions are introduced.
In the subsections the main components (global, not as an implementation) are introduced.
Outline of the thesis is presented.
\Cref{sec:Concepts} introduces the big parts of the puzzle.
\Cref{sec:Setup} gives details about the algorithms used.
\Cref{sec:Implementation} will give the implementation details.
\Cref{sec:Experiments} will show how data is gathered and shows the data.
\Cref{sec:Discussion} will discuss and relate the results
\Cref{sec:Conclusion} will conclude and give final remarks.

@@

\section{Concepts}
\label{sec:Concepts}
This chapter will introduce the abstract concepts which will be used.
There will be some formulas, but minimal implementations.
The focus is on the meaning of the terms and what they do and how they do it.
Real implementations will be discussed in \Cref{sec:Setup}.
